\title{Systematic Adaptive Control Design Using Sequential Loop Closure}
\author{Daniel Philip Wiese}
\prevdegrees{B.S., University of California, Davis (2011) \\
S.M., Massachusetts Institute of Technology (2013)}
\department{Department of Mechanical Engineering}
\degree{Doctor of Philosophy}
\degreemonth{June}
\degreeyear{2016}
\thesisdate{May 19, 2016}
\supervisor{Anuradha M. Annaswamy}{Senior Research Scientist}
\chairman{Rohan Abeyaratne}{Chairman, Department Committee on Graduate Students}
\maketitle

\cleardoublepage%
\setcounter{savepage}{\thepage}
\begin{abstractpage}
  \input{./chapters/abstract}
\end{abstractpage}

\cleardoublepage%

\section*{Acknowledgments}

I would like to first thank my advisor Dr.\ Anuradha Annaswamy for her guidance throughout this research and during my time at MIT.\@
Despite the pressure that often comes from ones advisor, working with Dr.\ Annaswamy has, overall, been a great experience.
While she was often very demanding, she was also very supportive, and mentored me to ensure that my worked was not lacking in rigor.
She worked hard to make sure my funding was available so that I was able to continue focusing on research.
She provided many opportunities to gain experience, including spending the summers doing internships with my research sponsors, or presenting lectures for her class.
She was always available to meet, but also flexible to accommodate other aspects of life.
For this and more I am thankful to have been able to complete my Ph.D. under Dr.\ Annaswamy.

I would also like to thank Dr.\ Jonathan Muse and Dr.\ Michael Bolender in the Aerospace Systems Directorate at the Air Force Research Laboratory for providing the vehicle model for this research, giving their support and mentorship along the way, and providing direction during my early summer work at AFRL.\@
They have both always been very helpful and accessible, and worked hard to ensure that they were able to provide the funding to support me throughout my doctoral studies.
Our interactions on a bi-weekly basis were crucial in keeping me focused, and their close involvement with me throughout my studies was greatly appreciated.
These discussions were invaluable to me as a developing researcher.
They were always helpful in providing feedback on my work, making sure that what I developed over the last several years was not only mathematically rigorous, but also practical.

Thanks to Dr.\ Eugene Lavretsky of Boeing Research \& Technology for contributing his precious advice and suggestions throughout this research.
My meetings with Eugene were some of the most productive I have ever had.
During these meetings he was able, in just moments, to parse out the real challenges I was facing, figure out what was really going on, and provide valuable insights and suggestions.
Working with Eugene helped me to learn how to distill and communicate complicated ideas, and focus on what really matters.
He was always incredibly positive and supportive, great fun to interact with, and was constantly sharing interesting problems to think about and work on.

I'd like to thank the other members of my committee for their involvement as well.
I took classes with both Professor How and Professor Hogan during my first semester as a master's student at MIT in the fall of 2011.
These classes were some of my most memorable experiences at MIT, and the way they communicated the ideas and material has stuck with me ever since.
I am grateful for agreeing to serve on my committee, the time they have given me, and the feedback and suggestions they have provided.

Thanks also to all of my lab mates for making the past years in lab enjoyable, and making our classes together fun.
Their assistance and friendship over these past years has been great.
It's been a pleasure to share an office with Max Zheng Qu and Heather Hussain, and it was always nice to take a break from work from time-to-time to chat.
Our time outside of the office with Ben Jenkins was always fun as well, providing the brain a much needed break from school.

I also want to thank Dr.\ Melanie Lutz, without whom I would not be where I am today.
She was the catalyst that, after several years of my failed attempts to become a successful student, jump-started me onto my path towards MIT and a Ph.D.
The three classes I took with her at Solano Community College during the fall semester of 2008 have turned out to be the most life-changing classes I have taken.
She was the most amazing classroom professor I have ever had, making the learning process engaging, stimulating, and challenging.
Over the years, she constantly pushed me to be my best, never allowing me to give less.
I want to thank her for all of the encouragement, inspiration, guidance, help, and friendship over the past eight years, and for being a truly outstanding mentor.

I would like to thank all of my friends for their patience and understanding over the last several years as well.
I appreciate my friends' support and encouragement throughout this journey, and putting up with my unavailability for months on end while I was busy with coursework and research.
I'd like to thank Alessandro Babini, with whom I am embarking on life's next big challenge.
Thank you for keeping that moving forward while allowing me to remain focused on my studies, and complete my PhD.

Finally, I would like to thank my Mom, Dad, and all of my family for their unconditional love, support, and understanding all these years.
They have kept me positive throughout the demanding years as a student, and always helped to keep me going.
Thanks for everything!

This research is funded by the Air Force Research Laboratory/Aerospace Systems Directorate grant FA 8650--07--2--3744 for the Michigan/AFRL Collaborative Center in Control Sciences, with Dr.\ Michael Bolender as technical monitor, and the Universal Technology Corporation (UTC) contract FA8650--10--D--3037 subcontract agreement 15--S2606--04--C19 for Adaptive Flight Control for Hypersonic Vehicles with Integrated Loops.
